---
title: 'Statistical report: N-terminal pro-B-type natriuretic peptide (NT-proBNP) in patients with heart failure, with a focus on worsening heart failure, in Sweden'
author: 'Statistician: Lina Benson'
  
date: "`r Sys.Date()`"
output:
  pdf_document:
    dev: cairo_pdf
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    number_sections: yes
link-citations: yes
bibliography: references.bib
nocite: '@*'
urlcolor: blue
linkcolor: black
header-includes:
   - \usepackage{subfig}
   - \usepackage{pdflscape}
   - \newcommand{\blandscape}{\begin{landscape}}
   - \newcommand{\elandscape}{\end{landscape}}
---

\newpage 
\tableofcontents 
\listoftables
\listoffigures
\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, include = TRUE, comment = "",
  warning = FALSE, message = FALSE, fig.pos = "H",
  fig.path = "../output/figs/"
)
options(knitr.kable.NA = "")
```

```{r adjust_directory_if_needed, include=FALSE}
# Uncomment lines below if rmd file is placed in a subdirectory
knitr::opts_knit$set(root.dir = normalizePath("../"))
```

```{r load_project}
# 1. Set options in config/global.dcf
# 2. Load packages listed in config/global.dcf
# 3. Import functions and code in lib directory

ProjectTemplate::reload.project()

cacheon <- TRUE
```             

# Data handling

## Data source

SCREAM2

## Inclusion/exclusion criteria

```{r flow}
default_kable(flow, caption = "Flowchart")
```

Note that patients who died in hospital are included in the analysis. Index date is defined as date of visit/discharge.  

First patient in: `r min(rsdata$shf_indexdtm)` and last patient in: `r max(rsdata$shf_indexdtm)`. 

The median age (IQR) is `r rsdata %>% summarise(med = fn(median(shf_age), dig = 1),
                                             q1 = fn(quantile(shf_age, probs = 0.25), dig = 1),
                                             q3 = fn(quantile(shf_age, probs = 0.75), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", q1, "-", q3, ")")) %>%
                                   pull(out)` and 
`r rsdata %>% count(shf_sex) %>%
  mutate(perc = fn(n / sum(n) * 100, 1)) %>%
  filter(shf_sex == "Female") %>%
  pull(perc)`% females.    

## Created variables 

The labvalues from SCREAM are selected within 14 days prior to index (including the index day) in SwedeHF, 
where the one closest to index is chosen. If there are two tests of the same lab 
from the same date the mean is calculated. 

```{r npr}
default_kable(cometa, caption = "Comorbidities and outcomes from VAL/KON")
```

```{r cci}
default_kable(ccimeta, caption = "Charlson Comorbidity index from VAL/KON")
```

Charlson Comorbidity index is calculated according to [@cci]. 

\newpage

# Statistical analysis 

## General

All analyses were performed using `r sessionInfo()$R.version$version.string` [@r]. 
The level of significance is set to 5%, two-sided. No adjustment for multiple 
comparisons were made and therefore the results should be viewed with some care.

## Missing data

Missing data was imputed with multiple imputation (n = 20) using mice [@mice]. 
Variables included in the model are indicated in 
Table \ref{tab:tab1}. 

## Baseline characteristics

```{r, child = "./src/tab1.Rmd"}

```

\clearpage

## Objective 1-2

### Trend of NT-proBNP testing after the index date

Patients are included in the denominator if their index date is prior to 1 Jan of the respective year and are alive and at risk on 31 Dec of the respective year. Only NT-proBNP testing AFTER index is considered. 

```{r, child = "./src/trend_nt.Rmd"}

```

\clearpage

### Frequencies and percentages of patients undergoing NT-proBNP within 30, 60 and 90 days before and after the index date 

Only patients at risk during the respective time periods are included in the analyses. 
Note that a patient (with 90 days follow-up) with one NT-proBNP testing at for example 25 days will be included in the numerator for all 3 post index periods. 

```{r, child = "./src/freq306090_nt.Rmd"}

```

\clearpage

### Number of days from the index date to the first post-diagnosis NT-proBNP test 

The number of days to NT-proBNP testing from day after index are presented in a Kaplan-Meier curve. 
Patients are censored at 2018-12-31 or death or move from Stockholm county. 

```{r, child = "./src/noday_nt.Rmd"}

```

\clearpage

\blandscape

### Number of NT-proBNP testing in the first 6 months, 1st and 2nd year after the index date per patient 

Only patients with complete follow-up during the specified time period are included. 

```{r, child = "./src/no612_nt.Rmd"}

```

\elandscape

```{r, child = "./src/no612_ntplots.Rmd"}

```

\clearpage

## Objective 3-4

### NT-proBNP concentration and distribution

```{r, child = "./src/conc_nt.Rmd"}

```

\clearpage

### Pattern of change 

Note that a cut-off of 3000 is used instead of 5000 (so stable low is defined as <3000). 

The 6 mo pre and post index are restricted to be within 6 months +/- 3 mo. The labvalues closest to 
6 months are selected. 

Note that there is no restriction that the patient needs to be alive at 6 months (since they need value at least 3 mo after index this is however indirectly applied).

```{r, child = "./src/change_nt.Rmd"}

```

\clearpage

## Objective 5:

### Association between change in NT-proBNP and WHFE at index

Only patients with measurements on NT-proBNP before and at index are included (no imputation performed for NT-proBNP). 
The change is calculated on the labvalue closest to 6 months prior to index. 

The association between change in NT-proBNP and WHFE is analysed using logistic regression models, 
partly crude including only change in NT-proBNP in the model 
and partly adjusted for variables indicated in Table \ref{tab:tab1}. The variables were selected based on clinical relevance.  

```{r, child = "./src/whfe_nt.Rmd"}

```

\clearpage

### Association between change groups (index to post) and risk of all-cause and CV deaths  

Only patients with measurements on NT-proBNP at and after index are included. 

The change is calculated on the labvalue closest to 6 months post index. 

The following outcomes are considered: 

- CV mortality
- All-cause mortality

Data were censored at 2018-12-31 or death/move from Stockholm county. Follow-up is started 6 mo after the indexdate. 

Time to first events were presented with Kaplan-Meier curves and Cox proportional hazards regressions were 
used to model the time to first event, partly crude including only change in NT-proBNP in the model 
and partly adjusted for 1) the index NT-proBNP value 2) the 6 mo NT-proBNP value. 

```{r, child = "./src/km.Rmd"}

```

\clearpage

```{r, child = "./src/out_nt.Rmd"}

```

\clearpage

The univariable cox regression including % change from index to 6 months 
modelled using a natural cubic spline with 4 degrees of freedom is depicted. The reference (HR = 1) is set at no change.

```{r, child = "./src/splines_out_nt.Rmd"}

```

\clearpage

## Objective 6

Association between testing of NT-proBNP and covariates was assessed using logistic regression. 
Variables < 0.1 in the crude models were selected for inclusion in the adjusted model. 

```{r, child = "./src/pred_nt.Rmd"}

```

\clearpage

# Reproducibility

## R session information {#sessioninfo}

```{r sessinfo}
sessionInfo()
```

## R code

The R code for all data handling and statistical analyses are found: 
https://github.com/KIHeartFailure/screamntprobnp. On publication
the repository will be made public so as to 
link to it from the resulting article for increased transparency and code sharing.
No data or output is stored in the repository. 

# References
